Histogram of Gradient Orientations of Signal Plots applied to P300 Detection

### 2017-12-21 response.letter.tex

2.1) the title is misleading since a first, and naive, read suggest being able to classify and characterize any EEG with their different components (i.e., P200, P300, P600, N100, N200, N400 to name a few). Clearly, the title does not match the content since the authors focus on the P300

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for the observation. The title of the article was changed accordingly to your suggestion.  
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

2.2) there is evidence of the benefits of working with these signals P300, but, is the proposed method extensible to other signals?, this is not clear from the manuscript.

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you very much for your comment.  We published in~\citep{Ramele2016} the application of the same method to identify rhythmic EEG events like Visual Occipital Alpha Waves and to classify Motor Imagery.  We are also working on unpublished material where we are analyzing the same approach for the detection of K-Complexes and classification of  SSVEP patterns but the results are in progress.  We added this information on the article (Introduction and Conclusion, section 3).  We hope this new information enriches the article and clarifies this point.

@@

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for the observation.  We modified all equations to emphasize the point that they are calculated in digital space, particularly on equation (4). We also added that image gradients are also calculated using finite differences, and added information about the support of each equation to explain that they are actually calculated in digital space.
With this, we hope we have full addressed your point.
 }
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

\subsection*{\ovalbox{Fourth Paragraph}}

4.1) Another issue that I find worrisome is the design of experiments. It seems that the number of subjects is rather low, it would be interesting to know why the authors decided to use such a subset.

\vskip+1ex
\noindent \dotfill
\vskip+1ex
%
\begin{quotation}
{\color{blue}
Thank you for your comment. The original dataset was published as the 008-2014 dataset at the BNCI-Horizon website, which is very important in BCI Research.  As far as we are aware of, experiments on this dataset of ALS patients has not been published by any independent group.
In order to address this issue, we included a new own dataset with the same experimental conditions that we generated ourselves, on healthy subjects.  We published this dataset on the Kaggle platform for replication.
}
\end{quotation}
%
\vskip+1ex
\noindent \dotfill
\vskip+1ex

### 2017-12-21 article.tex
Ultimos cambios finales. ya mandé el paper.

EEG signals are remarkably complex and have been characterized as a multichannel non-stationary stochastic process.  Additionally, they have high variability between different subjects and even between different moments for the same subject, requiring adaptive and co-adaptive calibration and learning procedures \citep{Clerc}.  Hence, this imposes an outstanding challenge that is necessary to overcome in order to extract information from raw EEG signals.

@@

\begin{equation}
 h(\theta,i,j) = 3s \sum_{\mathbf{p} \in B_{i,j}} w_\mathrm{ang}(\angle J(\mathbf{p}) - \theta)\, w_{ij}\left(\frac{\mathbf{p} - \mathbf{kp}}{3 s}\right)\, |J(\mathbf{p})|
\label{eq:histogram}
\end{equation}

@@

where $NN_T(d^l_u), \;l\in\{row,{col}\}$ is the set of the $k$ nearest neighbors to $d^l_u$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u)$.  This set is obtained according to the k-NBNN  algorithm~\citep{Boiman2008}.

@@

EEG data is collected in a single recording session. Participants are seated in a comfortable chair, with their vision aligned to a computer screen located one meter in front of them.  The handling and processing of the data and stimulations is conducted by the OpenVibe platform~\citep{Renard2010}.

@@

By analyzing the generated descriptors, a metric about the shape of the P300 response can also be derived.

El uso de descriptores basados en histogramas, presentado en este trabajo, también puede ser usado para derivar una métrica acerca de la forma de la señal de P300.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{This project was supported by the ITBACyT-15 funding program issued by ITBA University. The authors would like to thank Dr. Juliana Gambini for their insights and help with the detailed description of the histogram of gradients.}

>>>>

EEG signals are remarkably complex and have been characterized as a multichannel non-stationary stochastic process.  Additionally, they have high variability between different subjects and even between different moments for the same subject, requiring adaptive and co-adaptive calibration and learning procedures~\citep{Clerc}.  Hence, this imposes an outstanding challenge that is necessary to overcome in order to extract information from raw EEG signals.

@@

\begin{equation}
 h(\theta,i,j) = 3.s \sum_{\mathbf{p} \in B_{i,j}} w_\mathrm{ang}(\angle J(\mathbf{p}) - \theta)\, w_{ij}\left(\frac{\mathbf{p} - \mathbf{kp}}{3 s}\right)\, |J(\mathbf{p})|
\label{eq:histogram}
\end{equation}

@@

where $NN_T(d^l_u)$  is the set of the $k$ nearest neighbors to $d^l_u$ and $q$ is a template descriptor that belongs to $NN_T(d^l_u),\;l\in\{row,{col}\}$.  This set is obtained according to the k-NBNN  algorithm~\citep{Boiman2008}.

@@

EEG data is collected in a single recording session. Participants are seated in a comfortable chair, with their vision aligned to a computer screen located one meter in front of them.  The handling and processing of the data and stimuli is conducted by the OpenVibe platform~\citep{Renard2010}.

@@

%By analyzing the generated descriptors, a metric about the shape of the P300 response can also be derived.  
The use of descriptors based on  histogram of gradient orientation, presented in this work, also can be utilized for deriving a shape metric in the space of the P300 signals.
%El uso de descriptores basados en histogramas, presentado en este trabajo, también puede ser usado para derivar una métrica acerca de la forma de la señal de P300.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\acknowledgments{This project was supported by the ITBACyT-15 funding program issued by ITBA University.} %The authors would like to thank Dr. Juliana Gambini for their insights and help with the detailed description of the histogram of gradients.}
